Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-09 ms.tex
Finished subsection on rotations - short one

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em (ALL) Automated Lensing
  Learner}, a strong lensing identification pipeline that will be
publicly released as open source software.  In this first application,
we employ a fast feature extraction method, Histogram of Oriented
Gradients (HOG), to capture edge patterns that are characteristic of
strong gravitational arcs in galaxy-galaxy lensing.  We use logistic
regression to train a supervised classifier model on the HOG of HST-
and LSST-like images.  We use an area under the curve (AUC) of a
Receiver Operating Characteristic curve to assess model performance;
$AUC=1.0$ is an ideal trained classifier, and $AUC=0.5$ does not
provide any classification information.  Our best performing models on
a training set of 10,000 lens containing images and 10,000 non-lens
containing images exhibit an AUC of 0.975 for an HST-like sample.
However, for one orbit of LSST, our model only reaches an AUC of
0.625.  For 10-year mock LSST observations, the AUC improves to 0.809.
Model performance appears to continually improve with the size of the
training set.  Models trained on fewer images perform better in
absence of the lens galaxy, which can decrease the false positive rate
by blending with light from a lensed galaxy image near the central
lensing galaxy.  However, with larger training data sets, information
from the lensing galaxy improves model performance.  \wordchoice{Our
  results show efficient and effective methods for automated strong
  lensing identification with scalable linear methods that are easy to
  parallelize with existing open source tools and are reproducible on
  personal laptop computers.}
\end{abstract}
%-------------------------------------------------%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]\label{fig:hogimages}
\begin{center}
\includegraphics[width=2\columnwidth]{figures/tiles/compilations/hog_telescope_compilation.pdf}
\caption{Left to right show example image transforms of mock images
  from Figure~\ref{fig:mockimages} with a visualized histogram of
  oriented gradients.  The image transform picks up edge features,
  with arc features showing up as edges across radial orientations.}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%% \begin{figure*}[t]\label{fig:hogvisual}
%% \begin{center}
%% \includegraphics[width=1\columnwidth]{figures/105_lensed_imgs_1or5ppc32cpb1.pdf}\hspace{-20pt}
%% \includegraphics[width=1\columnwidth]{figures/105_lensed_imgs_1or9ppc16cpb1.pdf}\\
%% \includegraphics[width=1\columnwidth]{figures/lsst_105_lensed_imgs_1or4ppc3cpb3.pdf}\hspace{-20pt}
%% %\includegraphics[width=1\columnwidth]{figures/lsst_105_unlensed_imgs_1or4ppc3cpb3.pdf}\\
%% \includegraphics[width=1\columnwidth]{figures/lsst_stack_105_lensed_imgs_1or4ppc3cpb3.pdf}
%% %\includegraphics[width=1\columnwidth]{figures/lsst_stack_105_unlensed_imgs_1or4ppc3cpb3.pdf}
%% \caption{Top row: Visualized HOG feature vectors of the sample lensed
%%   HST-like image from Figure~\ref{fig:mockimages}.  Left: The
%%   corresponding HOG visualization with N$_\text{orient}=5$, (32, 32)
%%   pixels per cell, and (1, 1) cells per block.  Right: The HOG
%%   visualization of the same mock image with N$_\text{orient}=9$, (16,
%%   16) pixels per cell, and (1, 1) cells per block.  The brightest
%%   lines align with prominent edges.  Note, while fewer pixels per cell
%%   will better resolve the arc feature, the ``edge'' corresponding to
%%   the arc is less prominent.  Bottom row: Visualized HOG feature
%%   vectors for the LSST-like single epoch and stacked images with the
%%   best scoring HOG parameterization of N$_\text{orient}=4$, (3, 3)
%%   pixels per cell, and (3, 3) cells per block.  The single epoch image
%%   (left) does not have significant enough edge features to emerge in
%%   the HOG, but the stacked image (right) does.}
%% \end{center}
%% \end{figure*}
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\subsection{Effects of Dataset Size on Model Performance}\label{sec:datasize}
%-------------------------------------------------% 
%-------------------------------------------------% 
\subsubsection{Effects of Training Set Size}\label{sec:trainsetsize}
%-------------------------------------------------%

@@

%-------------------------------------------------% 
\subsubsection{Effects of Rotation on AUC}
%-------------------------------------------------% 
To augment our training set, we rotated each image in the set by
multiples of $90^o$.  Augmenting our data by a factor of four
naturally optimizes the use of available training data, since our
feature extraction method of HOG is not rotationally invariant.

We also tested the effects of testing our model on all four rotations
of the test set.  In Figure~\ref{fig:rotation_test}, we show the AUC
for different orientations of the datasets.  The x-axis corresponds to
each of our three datasets, \todo{with and without the lens galaxy}.
The y-axis shows the AUC.  The filled blue circles correspond to the
four AUCs calculated from test scores assigned to images at one
rotation.  The filled red stars correspond to the AUC calculated from
the average of all four test scores, which are systematically higher
than any one rotation.  The average score across all rotations for
each image is likely to be less noisy for the whole test sample,
giving an improved AUC.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:rotation_test}
\begin{center}
\includegraphics[width=\columnwidth]{figures/rotated_aucs/rotated_aucs.pdf}
\caption{Our summary figure: The AUCs of models trained on the full
  10,000 and tested on the holdout 1,000.  Blue circles: AUC
  calculated from scores of images at a given rotation (e.g. 0, 90,
  180, and 270 degrees).  Red stars: AUC calculated from the average
  score of all rotations of each image.  The average score produces an
  improved AUC in all data sets.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em (ALL) Automated Lensing
  Learner}, a strong lensing identification pipeline that will be
publicly released as open source software.  In this first application,
we employ a fast feature extraction method, Histogram of Oriented
Gradients (HOG), to capture edge patterns that are characteristic of
strong gravitational arcs in galaxy-galaxy lensing.  We use logistic
regression to train a supervised classifier model on the HOG of HST-
and LSST-like images.  We use an area under the curve (AUC) of a
Receiver Operating Characteristic curve to assess model performance;
$AUC=1.0$ is an ideal trained classifier, and $AUC=0.5$ does not
provide any classification information.  Our best performing models on
a training set of 10,000 lens containing images and 10,000 non-lens
containing images exhibit an AUC of 0.975 for an HST-like sample.
However, for one orbit of LSST, our model only reaches an AUC of
0.625.  For 10-year mock LSST observations, the AUC improves to 0.809.
Model performance appears to continually improve with the size of the
training set.  Models trained on fewer images perform better in
absence of the lens galaxy, which can decrease the false positive rate
by blending with light from a lensed galaxy image near the central
lensing galaxy.  However, with larger training data sets, information
from the lensing galaxy improves model performance.  \wording{Our
  results show efficient and effective methods for automated strong
  lensing identification with scalable linear methods that are easy to
  parallelize with existing open source tools and are reproducible on
  personal laptop computers.}
\end{abstract}
%-------------------------------------------------%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]\label{fig:hogimages}
\begin{center}
\includegraphics[width=2\columnwidth]{figures/tiles/compilations/hog_telescope_compilation.pdf}
\caption{Left to right show example image transforms of mock images
  from Figure~\ref{fig:mockimages} with a visualized histogram of
  oriented gradients.  The image transform picks up edge features,
  with arc features showing up as edges across radial orientations.
  \wording{Each of the oriented gradients within a cell is
    color-coded by magnitude, and represented as a line in the
    direction perpendicular to that gradient.}  The actual extracted
  features fill a one-dimensional feature vector comprised of the
  magnitudes of each of the oriented edges within the visualized
  cells.}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%% \caption{Top row: Visualized HOG feature vectors of the sample
%%   lensed HST-like image from Figure~\ref{fig:mockimages}.  Left:
%%   The corresponding HOG visualization with N$_\text{orient}=5$,
%%   (32, 32) pixels per cell, and (1, 1) cells per block.  Right: The
%%   HOG visualization of the same mock image with
%%   N$_\text{orient}=9$, (16, 16) pixels per cell, and (1, 1) cells
%%   per block.  The brightest lines align with prominent edges.
%%   Note, while fewer pixels per cell will better resolve the arc
%%   feature, the ``edge'' corresponding to the arc is less prominent.
%%   Bottom row: Visualized HOG feature vectors for the LSST-like
%%   single epoch and stacked images with the best scoring HOG
%%   parameterization of N$_\text{orient}=4$, (3, 3) pixels per cell,
%%   and (3, 3) cells per block.  The single epoch image (left) does
%%   not have significant enough edge features to emerge in the HOG,
%%   but the stacked image (right) does.}  \end{center} \end{figure*}
%%   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%-------------------------------------------------%
\subsection{Dataset Size Dependence}\label{sec:datasize}
%-------------------------------------------------% 
%-------------------------------------------------% 
\subsubsection{Effects of Training Set Size on Performance and Train Time}\label{sec:trainsetsize}
%-------------------------------------------------%

@@

%-------------------------------------------------% 
\subsubsection{Effects of Rotation on AUC}
%-------------------------------------------------%

% Augmenting dataset
To augment our training set, we rotated each image in the set by
multiples of $90^o$.  Augmenting our data by a factor of four
naturally optimizes the use of available training data, since our
feature extraction method of HOG is not rotationally invariant.  This
has an equivalent improvement to the study illustrated in
Figure~\ref{fig:trainsizeLSST10}.

% Rotations of the test set
We also tested the effects of evaluating our model on all four
rotations of the test set, and using the average score of each test
image to calculate the AUC.  In Figure~\ref{fig:rotation_test}, we
show the AUC for different orientations of the datasets.  The x-axis
corresponds to each of our three datasets, with and without the lens
galaxy.  The y-axis shows the AUC.  The filled blue circles correspond
to the four AUCs calculated when the model evaluated images at each of
the four rotations.  The filled red stars correspond to the AUC
calculated from the average of all four test scores, which are
systematically higher than any one rotation.  The average score across
all rotations for each image is likely to be less noisy for the whole
test sample, giving an improved AUC.

% Summary figure
Figure~\ref{fig:rotation_test} also summarizes the best-case results
of our models trained on our entire 10,000 training sets, and tested
on our holdout 1,000 test set.  Recall, however, that we expect model
performance on images containing the lens galaxy to improve further
with larger training sets (see Figure~\ref{fig:trainsizeLSST10}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/rotated_aucs/rotated_aucs.pdf}
\caption{Our summary figure: The AUCs of models trained on the full
  10,000 and tested on the holdout 1,000.  Blue circles: AUC
  calculated from scores of images at a given rotation (e.g. 0, 90,
  180, and 270 degrees).  Red stars: AUC calculated from the average
  score of all rotations of each image.  The average score produces an
  improved AUC in all data sets.  We expect the AUC to further improve
  with increased train size.}\label{fig:rotation_test}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
